Coursera "Practical Machine Learning" Project
========================================================
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 



## Data 

The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 


## Data Ingestion

First we download the training and test data sets and then we load them into R (training and testing data frames).

```{r}
# setwd("./R/PracticalMachineLearning/")
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",destfile="./data/pml-training.csv")
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",destfile="./data/pml-testing.csv")
training <- read.csv("./data/pml-training.csv")
testing <- read.csv("./data/pml-testing.csv")
```

# Feature Selection
In order to select the features to be used as predictors, we calculate the correlation between features and we locate those which have maximum impact on others (absolute value of correlation > 0.8)
The list of most important features is the following:

```{r}
M <- abs(cor(training[sapply(training, is.numeric)]))
diag(M) <- 0
which(M > 0.8,arr.ind=T)
```

## Preparation for Cross-Validation

In order to measure our model we partition our training set into two parts (75%, 25%): train1, test1 data frames.

```{r}
library(caret)
library(randomForest)
inTrain <- createDataPartition(training$classe,p=0.75,list=FALSE)
train1 <- training[inTrain,]
test1 <- training[-inTrain,]
```


## Prediction Models

We will be using Random Forest based on the first part of our training set (train1). 

```{r}
library(caret)

rf1 <- train(classe ~ user_name+roll_belt+pitch_belt+yaw_belt+gyros_belt_x+gyros_belt_y+gyros_belt_z+accel_belt_x+accel_belt_y+accel_belt_z+magnet_belt_x+magnet_belt_y+magnet_belt_z+roll_arm+pitch_arm+yaw_arm+total_accel_arm+gyros_arm_x+gyros_arm_y+gyros_arm_z+accel_arm_x+accel_arm_y+accel_arm_z+magnet_arm_x+magnet_arm_y+magnet_arm_z,train1,method="rf")
```

## Error Estimation

We will estimate errors by comparing predictions towards the real classe values of the second part of training set (test1).
Error probability and number of correct predictions are displayed below:

```{r}
pred1 <- predict(rf1,test1[,-160])
A <- table(pred1,test1$classe)
error1 <- (sum(A)-sum(diag(A)))/sum(A)
predRight1 <- pred1==test1$classe

error1
sum(predRight1)
```


## Final Prediction Outcome

The prediction outcome of the final model on the original test set is:

```{r}
predict(rf1,testing)
```

